Dynamics of international giving: Exploring individual donor preferences

### 2020-06-17 manuscript/manuscript.Rmd
My R&R changes

```{r setup, include=FALSE}
library(tidyverse)
library(scales)
library(glue)
library(here)

@@

Over the last three decades, governments across the world have sought to limit the work of nonprofits - a phenomenon knowns as `closing civic space' in the policy arena [@Carothers:2015; @DPR:2015]. Government have enacted laws creating barriers to advocacy, funding, and entry, adversely affecting the work of thousands of NGOs, which often receive a large amount of resources from donors across the world. This is particularly the cause in countries with weak institutions and poor governance, where direct aid transfers to governments raise issues of misuse and bureaucratic inefficiency and donors seek out NGOs for projects (Dietrich 2013). Consequently, official aid flows to NGOs have decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. While we know there has been a negative impact on public funding flows, we know relatively little about how this crackdown affects private donors, especially individual donors' willingness to donate.

Charitable-giving to nonprofits working in international affairs has been steadily increasing over the past two decades, with a noticeable increase over the last decade. Between 2014 and 2016, giving to international affairs increased by 14.1% [@givingusa2017report]. This amount is not insignificant—by 2016, total private giving to civil society organizations working in international affairs had increased to \$22.03 billion [@givingusa:2017]. Most notably, a large portion of this is driven by individuals making small-scale donations. In 2017, individuals gave $286.7 billion - or approximately 70\% of total giving to international affairs. This further increased to $427.71 billion in 2018 - 85.7% of which was through individuals [givingusa:2019]. Given that a majority of this research is restricted to elite, high net-worth donors [@USTrust2014], these figures show that it is imperative to understand motivations and attributes of individual donors. What factors affect individual donor preferences on giving to international causes? More specifically, how does information about regulatory environments abroad, as well as organizational-level attributes of nonprofits, such as their main focus areas and sources of funding affect donors' willingness to donate internationally?

Given information asymmetry and time constraints, individual donors often use heuristics to simply their decision-making [@TverskyKahneman:1974; @CrosonShang:2011; @Tremblay-BoirePrakash:2017], We leverage an online survey experiment using Amazon's Mechanical Turk to assesss how different structural and organizational-level heuristics affect individual donors' willingness to donate. We find that learning about more repressive NGO regulatory environments increases generosity in that already likely donors are willing to donate substantially more to legally restricted NGOs.

This generosity persists when mediated by two important organization-level heuristics: NGO issue area and its main funding sources. Learning about crackdown increases individuals' willingness to donate to privately funded nonprofits. This may be because the survival of the NGO appears to be in question and the donor's assessment might be that their donation actually makes a difference. These results are particularly substantive for privately-funded human rights NGOs facing crackdown - with donors showing an increased willingness to not just donate to them, but also to donate more to them. While not unsurprising, this does suggest that private donors to human rights NGOs likely know that their work is always challenging to governments, and that such organizations need greater support when facing stricter regulatory environments.

Our study makes a significant contribution to the scholarship and practice on dynamics of individual-giving and indicates the promise and limits of different framing appeals by nonprofits on individuals' willingness to donate. First, while factors giving shaping individual-giving to domestic causes have been thoroughly explored in existing literature, the dynamics of internatoinal giving are under-examined in the U.S. literature on individual donor preferences.[^exception] It is even more essential to understand as philanthropy to international causes can be harder to motivate because of the number of recipients are larger and far removed from the donor [@CasaleBaumann:2015, 100].

Second, most nonprofits working internationally have traditionally relied on government and foundation funding. But in the era of closing civic space, these groups may need to tailor their fundraising strategies to individual donors. Framing, or the process through which actors present information to infleunce perceptions of behavior, is particular important in this regard.[^framing] However, we lack systematic studies of which frames may be effective when making appeals to donors considering international charitable-giving. Our study looks at the impact of political frames such as regulatory crackdown, and how it interacts with organizational attributes to signal NGO deservingness. In doing so, it can help nonprofits frame their appeals.

CHANGE
Below, we summarize existing research on the determinants of giving to NGOs and lay out our expectations regarding how crackdowns on NGOs may affect private philanthropy. We then describe our survey and present the results. We conclude with implications for NGOs working in repressive countries and lay out questions ripe for future research.

#Philanthropy towards international nonprofits

Thousands of international NGOs in the Global South receive funds from a variety of public and private sources each year.[^NGO] Though our focus in this paper is one private funding, it is essential to differentiate between the two sources. Public aid (more commonly known as traditional donor aid) channels funds from official aid agencies towards nonprofits through a variety of bilateral and multilateral institutions. Private funds, which are comparatively newer international donors compared to traditional donor aid, can come from foundations, private corporations and individuals.

In this paper, we specifically look at individual-level determinants to foreign giving. As described in the introduction, this is because [ADD PART ABOUT ANY PUBLIC FIGURES AND HOW IT DECLINING] private-giving to nonprofits working in international affairs has been consistently increasing over this past decade.[^internationalaffairs] Moreover, private giving by individuals has far outstripped giving by private foundations - in 2018, charitable-giving by individuals amounted to $292.09 billion, compared by $75.86 billion by foundations and $20.05 billion by corporations. Thus, giving by foundation represented 18\% of all charitable-giving in 2018, which was not only lesser than the proportion of giving by private individuals, but also giving by Bill and Melinda Gates Foundation accounted for more than half of it [givingusa:2019].

Given these substantive amounts, understanding individual-level determinants of giving to international nonprofits is important to investigate as individual private donors may not necessarily mirror public donor choices. While there is evidence that citizens generally support the goals of aid agencies (Milner and Tingley 2010), surveys show that the American public is notoriously misinformed about US foreign aid [@norris:2017]. Further, foreign aid and democracy aid is often considered unresponsive to developments on the ground [@brainard2007, @Carothers:2015] and public aid decisions - especially on final recipients are not easily accessible. There may also be heterogeneity within how donor governments allocate funds. Emerging donor government differ from established donors in that they demand fewer human rights, governance or environmental conditions preferred by Western donors [@fenglerkharas].

More recent research on the allocation of private aid flows has also found little support for the idea that individual donor behavior mimics that of official donors. Desai and Kharas (2018) find that unlike official donors  whotend to channel funds to countries with better institutional quality (control of corruption) and as a performance reward (Bermeo 2011, Dietrich ad Wright 2017), individual donors do not use the same performance-based metrics. Rather, individuals are more inclined to donate to groups in countries experiencing a humanitarian crisis - in particular, those driven by natural disasters. They go on to conclude that compared to donor agencies, "private donors respond to different project and country characteristics." (517) Given these above reasons, decision-making by the average individual donor may therefore not have the same logic as donor governments and should be investigated on its own terms.

Understanding motivations of individual private donors is also essential because private funding flows towards nonprofits working internationally appear to not be as adversely impacted as public funding flows in the face of the global crackdown on NGOs. Prior research has demonstrated that restrictions on cross-border funding has reduced official aid and official development assistance to repressive countries (CITE). But data from the Foundation Center shows no correlation between amount of funding flows from US private foundations to a given country and that country's score on the Index of Philanthropic Freedom. While this index is comparatively new and data is available only 2015 onwards (CHECK), preliminary evidence shows that, "US foundations remain committed to channeling funds to countries where the legal environment is growing increasingly unfriendly." As such, it is imperative to understand whether individual donors mirror the trajectory of foundations when giving internationally, especially to countries with restrictive legal environments.

##Key drivers of individual-level philanthropy

Substantial research on charitable giving has looked at the motivations of private donors, with much of this work focusing on the individual-level determinants of giving [@Wiepking:2010; @BekkersWiepking:2011]. However, research on individual donor behavior has overwhelming examined giving to local—rather than international—causes. This literature has primarily focused on the importance of X main factors, especially when considering charitable behavior towards international causes: (1) the role of social capital and social networks; (2) the role of individual experiences such as higher education and levels of religiosity; and (3) third party certifications.

One major area of inquiry in philanthropic studies looks at the role of social capital, or the "networks, norms, and social trust that facilitate coordination and cooperation for mutual benefit." (Putnam 1997, 67). While quite broad, the literature further concentrates on a specific kind of social capital-associational capital-which refers to, "the networks of formal and informal relationships to which people are associated." [@havens2007, 240]. Prior research has found that those who participate in a variety of associations are more likely to make larger donations [@hossainlamb2017]. Further, those who volunteer their time (compared to those who did not), were both more likely to make charitable donations and make bigger donations. This is because volunteering exposes individuals to the need for donations to charities and nonprofits and made them feel more compelled to do so [@hossainlamb2012].

The role of social networks is also important in this regard. Looking at relational configurations in social networks, Herzog and Yang find that having both a giver and a solicitor in one's social network increase the likelihood of that individual's charitable-giving.

Individual experiences such as access to higher education and increased religiosity can also impact philanthropic giving. Higher education fosters prosocial motivations and also brings people into particular social networks that entail a higher level of solicitations [@BekkersWiepking:2011]. This in turn has been shown to be particularly vital in giving to international causes. Additional education has also been shown to increase abstract thinking [@wiepkingmass, 1978]. This is particular important for donating to international relief organizations than donating to domestic organizations [@micklewright2009].

Levels of religiosity are also a significant predictor of the likelihood and level of giving. People who regularly attention religious services are more likely to have prosocial values, face a stronger norm of charitable-giving, and may feel pressure to conform that group and community standard [@wiepking2010]. Overall, individuals with higher levels of income, education, and greater religious proclivities have been shown to not only be more likely to give internationally, but also to give higher amounts [@Rajan:2009; @CasaleBaumann:2015].

Finally, information on organizational characteristics and third party certifications can also impact levels of individual charitable-giving, through the results are mixed. Existing research shows that information on the efficiency of NGOs merely steers rather than encourages or discourages overall donations [@ryazanov:efficiency]. The effects of changes in charity watchdog ratings and nonprofit accountability systems on individual donors is also mixed, showing that donors do not have the time to research charities thoroughly [@Bekkers:2010] and that ratings by charity watchdogs do not really affect donor support for these nonprofits (Szper and Prakash 2011).

While a robust literature as shown above looks at the drivers of individual-level philanthropy, it overwhelmingly focusses on giving to local rather than international causes. This focus on local or domestic causes means that that the precise causal mechanisms through which cross-border charity can be motivated and mobilized is still under-examined (Desai and Kharas, 517). This problem is further compounded by the fact that there is greater paucity of data and diversity of sources regarding international-giving, which further complicates a systematic examination.

Indeed, most research on individual-giving to international causes has been restricted to elite, high net-worth donors [@USTrust2014].[^highnetworth] And recent research has on the average donor finds that, "donors are more likely to donate to a charity operating locally than to a charity providing identical service abroad." (TB Prakash 2017, 644) And as explained above in section X, private donors do not necessarily mirror the choices made by public donors and that individuals rely on signals and heuristics that may differ from the benchmarks used by large governmental and private donors.

These signals and heuristics, we argue, can be even more important in giving to international causes. In theory, donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation, but this rarely happens in practice [@Tremblay-BoirePrakash:2017]. Instead, donors respond to a host of heuristics and signals when deciding whether to donate to an NGO, since seeking complete information about an organization's deservingness and efficiency is costly and time-consuming. These heuristics are commonly rooted in donor perceptions of organizational characteristics—donors make cursory judgments about an organization's issue area, mission, vision, and values, and seek out supplementary information from friends, family, and acquaintances [@SzperPrakash:2011; @Sloan:2009]. Nonprofits fundraising internationally may be therefore more reliant on framing. Below, we theorize the effect of different frames surrounding a country's legal environment and nonprofit organizational characteristics on individual donor preferences.

#Theorizing the effects of political and organizational frames on individual-giving

@@

The *contentiousness of NGO issue areas*—or the degree to which an NGO's programming is compatible with government preferences, can also serve as an organization-level heuristic.[^targeting] NGOs that address humanitarian issues such as relief and development have broader appeal to donors and these issues rarely challenge the government. Further, their programmatic output is easily quantifiable and donors can see the results of their charity more readily [@Bush:2015]. Previous research has shown that that public donors do indeed differentiate between NGO issue areas, especially when these organizations are working in repressive contexts. In particular, the Organisation for Economic Cooperation and Development (OECD) donors respond to repressive NGO environments, especially those containing barriers to advocacy, by decreasing funds for groups working on politically sensitive causes such as anti-corruption initiatives, elections, human rights, legal reform and security sector reform, and instead increasing funding for groups working on relatively tamer causes such as health, education and agriculture [CITE US].

Issue area contentiousness directly influences legal crackdowns, however, and confounds the heuristic effect of crackdowns on donor behavior (see @fig:causal-path). NGOs working on more contentious human rights issues can be challenging to regimes [@Heiss:2017]. These organizations are often viewed as politically motivated outsiders and run a higher risk of getting expelled [@DPR:2015], increasing the possibility of wasting donor resources. There is evidence that this a concern for private donors. In a survey of women’s NGOs across the globe conducted by Mama Cash and Urgent Action Fund, many NGOs argued that private donors (especially foundations) were withdrawing from funding sensitive issues in repressive contexts, which in turn was hurting women and trans rights organizations. Some private funders, these groups argued, were also more likely to fund well-established groups that were more likely to survive crackdown.

However, compared to our existing knowledge on public donors and foundation donors, we lack knowledge about how individual donors respond to frames emphasizing one nonprofit issue area over another, especially a contentious area over a tamer one. NGOs working on tamer issues such as health, education, disaster relief, and humanitarian aid can often frame their activities as in need of greater support—the perception of need, whether through personal experiences or media framing. This may strongly influence individual decisions to donate to relief efforts [@BekkersWiepking:2010, 6–8]. Individual donors may believe that it is more difficult to improve human rights in the host country and subsequently shy away from donating to NGOs working on contentious causes.

@@

The final heuristic we examine is the *source of NGO funding*. Government grants to nonprofits can often crowd out private funding, though evidence is mixed on whether this completely or partially crowds out private dollars [@Steinberg:1991; @Heutel:2014]. Donors may feel less inclined to contribute to INGOs receiving funds from agencies like USAID because such funding can signal a lack of independence from the government. Organizations that receive substantial government funding frequently avoid programming that would question the donor government even if such programming is in line with their mission [@Stroup:2012]. In contrast, individual donors who see that organizations are privately funded may feel that they can also contribute and help.

@@

> **H~3a~**: If donors learn that legally besieged NGOs do not receive substantial funding from government sources, then individual donors will be more likely to donate to them.
>
> **H~3b~**: If donors learn that legally besieged NGOs do not receive substantial funding from government sources, then individual donors will donate more to them.

Importantly, when testing these hypotheses, we do not assume that individual donors have perfect knowledge about an NGO or the country an NGO operates in. Donors do not always know about anti-NGO laws in host countries or who funds an NGO when they make their contributions. Moreover, the average donor will not know exactly what "legal crackdown" entails for any given organization (i.e. does it mean an NGO was expelled, or that its assets were frozen, or that it received a fine?). For the sake of this experiment, this ambiguous definition is not a central concern as the heuristics we explore here are explicitly simplified shortcuts to donor information-gathering. Knowing that an organization works with human rights issues (even if donors are not familiar with specific kinds of rights) should be enough to trigger the issue-based deservingess heuristic. Similarly, knowing that an NGO faces legal trouble should be enough to trigger the structural deservingness heuristic.

@@

To measure the effect of structural and organization-level heuristics on individual donor preferences, we use an experiment to vary the different signals donors are exposed to. Comparing these effects allows us to measure the relative strength of these heuristics, and ultimately help NGOs tailor their fundraising strategies for private individuals.

@@

We test these hypotheses with a vignette-based survey experiment fielded through Amazon's Mechanical Turk.[^mturk] Our target population is the portion of Americans hypothetically willing to donate money for human rights and humanitarian work abroad. Our convenience sample (March 2018, N = `r nrow(results)`) generally approximates the characteristics of our target population, since it is younger, more educated, wealthier, and more likely to donate to charities than nationally representative samples.[^supp]

@@

We presented participants with a short paragraph with three manipulated frames, each highlighting different factors that might influence patterns of private philanthropy: (1) crackdown/no crackdown, (2) humanitarian assistance/human rights, and (3) government/private funding. We used the International Rescue Committee (IRC) as our example NGO because its work easily applies to each possible frame. IRC provides humanitarian assistance and engages in human rights advocacy for refugees, more than a quarter of IRC's income comes from both government grants and private donations, and it works in countries with and without anti-NGO laws. We use a 2×2×2 between-subject factorial design with participants randomly assigned to one of eight versions of the following vignette:

@@

We collected two outcome measures: (1) how likely participants would be to donate to IRC (measured with a 5-point scale ranging from "Extremely likely" to "Extremely unlikely"), and (2) how much participants would hypothetically donate to IRC if they had an extra $100. We collapsed the likelihood scale into a binary variable measuring whether the participant is likely ("Extremely likely" and "Somewhat likely") or not likely (all other responses) to donate.[^collapse] We also included an exploratory free response question asking respondents to justify their choice.

@@

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(pander)
library(scales)
library(glue)
library(here)

source(here("analysis", "options.R"))

@@

Over the last three decades, governments across the world have sought to limit the work of nonprofits and nongovernmental organizations (NGOs)—a phenomenon commonly known as "closing civic space" [@Carothers:2015; @DPR:2015]. Repressive governments have enacted laws creating barriers to advocacy, funding, and entry, adversely affecting the work of thousands of NGOs, which typically rely on funding from donors across the world. This is particularly the case in countries with weak institutions and poor governance, where international donors typically seek out NGOs to implement their projects, as direct aid transfers to recipient governments pose the risk of misuse and bureaucratic inefficiency [@Dietrich2013]. As a result of this global crackdown on civil society, however, official aid flows to NGOs have decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. While we know there has been a negative impact on funding from official governmental donors, we know relatively little about how this crackdown affects private donors, in particular how it influences individual donors' willingness to engage in philanthropy.

Charitable giving to nonprofits working in international affairs has steadily increased over the past two decades, with a noticeable increase over the last decade. Between 2014 and 2016, giving to international affairs-focused issues increased by 14.1% [@givingusa2017report]. This amount is not insignificant—by 2016, total private giving to civil society organizations working in international affairs had increased to \$22.03 billion [@givingusa:2017]. Most notably, a large portion of this growth is driven by individuals making small-scale donations. In 2017, individuals gave $286.7 billion, or approximately 70% of total giving to international affairs. This further increased to $427.71 billion in 2018, 85.7% of which was through individuals [givingusa:2019]. Much of the existing research on donations to international causes is restricted to elite, high net-worth donors [@USTrust2014], but these large amounts demonstrate that it is imperative to understand the motivations and attributes of more ordinary individual donors. What factors affect individual donor preferences on giving to international causes? More specifically, how does information about regulatory environments abroad, as well as organizational-level attributes of NGOS— such as their main issues of focus and sources of funding—affect donors' willingness to donate internationally?

Given information asymmetry and time constraints, individual donors often use heuristics to simply their decision-making [@TverskyKahneman:1974; @CrosonShang:2011; @Tremblay-BoirePrakash:2017], We leverage an online survey experiment using Amazon's Mechanical Turk to assess how different structural and organizational-level heuristics affect individual donors' willingness to donate. We find that learning about more repressive NGO regulatory environments increases generosity in that already likely donors are willing to donate substantially more to legally restricted NGOs.

This generosity persists when mediated by two important organization-level heuristics: NGO issue area and funding sources. Learning about crackdown increases individuals' willingness to donate to privately funded nonprofits. This may be because the survival of the NGO appears to be in question and the donor's assessment might be that their donation actually makes a difference. These results are particularly substantive for privately-funded human rights NGOs facing crackdown, with donors showing an increased willingness to not just donate to them, but also to donate more to them. While not unsurprising, this does suggest that private donors to human rights NGOs likely know that their work is always challenging to host governments, and that such organizations need greater support when facing stricter regulatory environments.

Our study makes a significant contribution to research on individual giving and indicates the promise and limits of different framing appeals on individuals' willingness to donate. First, while factors shaping individual giving to domestic causes have been thoroughly explored in existing literature, the dynamics of international giving and individual donor preferences are less explored.[^exceptions] Understanding the factors driving individual giving is particularly important when when looking at international issues, as philanthropy to international causes can be harder to motivate because of the number of recipients is larger and further removed from the donor [@CasaleBaumann:2015, 100].

Second, most nonprofits working internationally have traditionally relied on government and foundation funding. But in the era of closing civic space, these groups may need to tailor their fundraising strategies to individual donors. Framing, or the process through which actors present information to influence perceptions of behavior, is particularly important in this regard.[^framing] However, we lack systematic studies of which frames may be effective when making appeals to donors considering international philanthropy. Our study examines the impact of political frames such as regulatory crackdown, and how these frames interact with organizational attributes to signal NGO deservingness. Accordingly, our results can help international nonprofits frame their appeals for funding, especially when facing restricted legal space abroad.

Below, we summarize existing research on the determinants of giving to NGOs and lay out our expectations regarding how crackdowns on NGOs may affect how individuals engage in philanthropy. We then describe our survey experiment and present the results. We conclude with implications for NGOs working in repressive countries and lay out questions ripe for future research.

# Philanthropy towards international NGOs

Thousands of international NGOs in the Global South receive funds from a variety of public and private sources each year.[^NGO] Though our focus in this paper is one private funding, it is essential to differentiate between the two sources. Public aid (more commonly known as traditional donor aid) channels funds from official aid agencies towards nonprofits through a variety of bilateral and multilateral institutions. Private funds—which are comparatively newer international donors compared to traditional donor aid—can come from foundations, private corporations, and individuals.

In this paper, we specifically look at individual-level determinants to international philanthropy. Private giving to nonprofits working in international affairs has increased consistently over the past decade.[^internationalaffairs] Moreover, donations from individuals have far outstripped foundation and corporate giving—in 2018, individual donations amounted to \$292.09 billion, compared to \$75.86 billion from foundations and \$20.05 billion by corporations. While foundations are often seen as the primary funders of international charities, donations from foundations accounted for only 18% of all charitable giving in 2018—and grants from the Bill and Melinda Gates Foundation accounted for more than half of all foundation giving [@givingusa:2019]. Individual donors play an incredibly important—and underexplored—role in funding international NGOs.

Given these substantive amounts, understanding individual-level determinants of giving to international nonprofits is important to investigate as individual private donors may not necessarily mirror the preferences of official aid agencies. While there is evidence that citizens generally support the goals of aid agencies [@MilnerTingley2010], surveys show that the American public is notoriously misinformed about US foreign aid [@norris:2017]. Further, both foreign aid and democracy aid are often considered unresponsive to political developments in recipient countries [@brainard2007; @Carothers:2015] and public aid decisions—especially about final recipients and amounts—are not easily accessible. There may also be heterogeneity within how donor governments allocate funds. For instance, unlike established donors like the United States and other OECD countries, emerging donor governments tend to demand fewer human rights, governance, or environmental preconditions [@fenglerkharas].

More recent research on the allocation of private aid flows has also found little support for the idea that individual donor behavior mimics that of official donors. @desaikharas find that unlike official donors who tend to channel funds to countries with better institutional quality or as a reward for governmental performance [@Bermeo:2011; @DietrichWright:2015], individual donors do not use the same performance-based metrics. Rather, individuals are more inclined to donate to groups in countries experiencing a humanitarian crisis—in particular, crises following natural disasters. @desaikharas conclude that compared to donor agencies, "private donors respond to different project and country characteristics" (p. 517) Given these above reasons, the average individual donor will not follow the same decision-making logic as donor governments, and individual international philanthropy should therefore be investigated on its own terms.

Understanding the motivations of individual private donors is also essential because private funding flows towards nonprofits working internationally appear to not be as adversely impacted as public funding flows in the face of the global crackdown on NGOs. Prior research has demonstrated that restrictions on cross-border funding has reduced official aid and official development assistance to repressive countries [@citeme]. But data from the Foundation Center shows no correlation between amount of funding flows from US private foundations to a given country and that country's score on the Index of Philanthropic Freedom. Additionally, preliminary evidence shows that "US foundations remain committed to channeling funds to countries where the legal environment is growing increasingly unfriendly." [@citeme-too]. As such, it is imperative to understand whether individual donors mirror the trajectory of foundations when giving internationally, especially to countries with restrictive legal environments.

## Key drivers of individual-level philanthropy

Substantial research on charitable giving has looked at the motivations of private donors, with much of this work focusing on the individual-level determinants of giving [@Wiepking:2010; @BekkersWiepking:2011]. However, research on individual donor behavior has overwhelming examined giving to local—rather than international—causes. This literature has primarily focused on the importance of three main factors, especially when considering charitable behavior towards international causes: (1) the role of social capital and social networks; (2) the role of individual experiences such as higher education and levels of religiosity; and (3) third party certifications.

One major area of inquiry in philanthropic studies looks at the role of social capital, or the "networks, norms, and social trust that facilitate coordination and cooperation for mutual benefit" [@Putnam:1995, 67]. While quite broad, the literature concentrates on "the networks of formal and informal relationships to which people are associated," [@havens2007, 240] or associational capital. Prior research has found that those who participate in a variety of associations are more likely to make larger donations [@hossainlamb2017]. Further, those who volunteer their time (compared to those who did not), were both more likely to make charitable donations and make bigger donations, since volunteering exposed individuals to the need for charitable donations and made philanthropy more compelling [@hossainlamb2012]. An individual's social networks also play an important in deciding to donate. Looking at relational configurations in social networks, @HerzogYang:2018 find that having both a giver and a solicitor in one's social network increase the likelihood of that individual's charitable giving.

Individual experiences such as access to higher education and increased religiosity can also influence philanthropic giving. Higher education fosters prosocial motivations and brings people into particular social networks that entail a higher level of solicitations [@BekkersWiepking:2011]. Education also increases abstract thinking [@wiepkingmaas, 1978], which is particularly important for donating to more distant international relief organizations than donating to local, more visible nonprofits [@Micklewright:2009]. Levels of religiosity are also a significant predictor of the likelihood and level of giving. People who regularly attend religious services are more likely to have prosocial values, face a stronger norm of charitable giving, and may feel pressure to conform to group and community standards [@wiepking2010]. Overall, individuals with higher levels of income, education, and greater religious proclivities have been shown to not only be more likely to give internationally, but also to give higher amounts [@Rajan:2009; @CasaleBaumann:2015].

Finally, information about organizational characteristics and third party certifications can also impact levels of individual philanthropy, through results are mixed. Existing research shows that information on the efficiency of NGOs merely steers rather than encourages or discourages overall donations [@ryazanov:efficiency]. The effects of changes in charity watchdog ratings and nonprofit accountability systems on individual donors is also mixed, showing that donors do not have the time to research charities thoroughly [@Bekkers:2010] and that ratings by charity watchdogs do not really affect donor support for these nonprofits [@SzperPrakash:2011].

While this robust literature looks at the drivers of individual-level philanthropy, it overwhelmingly focusses on giving to local rather than international causes. This domestic focus is warranted, though: on average, "donors are more likely to donate to a charity operating locally than to a charity providing identical service abroad" [@Tremblay-BoirePrakash:2017, 644]. However, this focus on domestic philanthropy causes means that that the precise causal mechanisms through which cross-border charity can be motivated and mobilized is still under-examined [@desaikharas, 517]. This problem is further compounded by the fact that there is greater paucity of data and diversity of sources regarding international philanthropy. Indeed, most research on individual-giving to international causes has been restricted to elite, high net-worth donors [@USTrust2014].[^highnetworth] Given the increasing importance of more ordinary individual donations for international NGOs, it is crucial to better understand what motivates private donors to engage in internationally-focused philanthropy.

As explored previously, individual private donors do not necessarily mirror the choices made by public donors. Individuals rely on signals and heuristics that differ from the benchmarks used by large governmental agencies and private foundations. These signals and heuristics, we argue, can be even more important in giving to international causes. In theory, donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation, but this rarely happens in practice [@Tremblay-BoirePrakash:2017]. Instead, donors respond to a host of heuristics and signals when deciding whether to donate to an NGO, since seeking complete information about an organization's deservingness and efficiency is costly and time-consuming. These heuristics are commonly rooted in donor perceptions of organizational characteristics—donors make cursory judgments about an organization's issue area, mission, vision, and values, and seek out supplementary information from friends, family, and acquaintances [@SzperPrakash:2011; @Sloan:2009]. Nonprofits fundraising internationally may be therefore more reliant on framing. Below, we theorize the effect of different frames surrounding a country's legal environment and nonprofit organizational characteristics on individual donor preferences.

# Theorizing the effects of political and organizational frames on individual giving

@@

The *contentiousness of NGO issue areas*—or the degree to which an NGO's programming is compatible with government preferences, can also serve as an organization-level heuristic.[^targeting] NGOs that address humanitarian issues such as relief and development have broader appeal to donors and these issues rarely challenge the government. Further, their programmatic output is easily quantifiable and donors can see the results of their charity more readily [@Bush:2015]. Previous research has shown that that public donors do indeed differentiate between NGO issue areas, especially when these organizations are working in repressive contexts. In particular, donor agencies in the Organisation for Economic Cooperation and Development (OECD) respond to repressive NGO environments, especially those containing barriers to advocacy, by decreasing funds for groups working on politically sensitive causes such as anti-corruption initiatives, elections, human rights, legal reform and security sector reform, and instead increasing funding for groups working on relatively tamer causes such as health, education and agriculture [@ChaudhryHeiss:2018].

Issue area contentiousness directly influences legal crackdowns, however, and confounds the heuristic effect of crackdowns on donor behavior (see @fig:causal-path). NGOs working on more contentious human rights issues can be challenging to regimes [@Heiss:2017]. These organizations are often viewed as politically motivated outsiders and run a higher risk of getting expelled [@DPR:2015], increasing the possibility of wasting donor resources. There is evidence that this a concern for private donors. In a survey of women’s NGOs across the globe conducted by Mama Cash and Urgent Action Fund, many NGOs argued that private donors (especially foundations) were withdrawing from funding sensitive issues in repressive contexts, which in turn hurt women and trans rights organizations. Some private funders, these groups argued, were also more likely to fund well-established groups that were more likely to survive crackdown.

However, compared to our existing knowledge on public donors and foundation donors, we lack knowledge about how individual donors respond to frames emphasizing one nonprofit issue area over another, especially a contentious area over a tamer one. NGOs working on tamer issues such as health, education, disaster relief, and humanitarian aid can often frame their activities as in need of greater support—the perception of need, whether through personal experiences or media framing. This may strongly influence individual decisions to donate to relief efforts [@wiepking2010]. Individual donors may believe that it is more difficult to improve human rights in the host country and subsequently shy away from donating to NGOs working on contentious causes.

@@

The final heuristic we examine is the *source of NGO funding*. Government grants to nonprofits can often crowd out private funding, though evidence is mixed on whether this completely or partially crowds out private dollars [@Steinberg:1991; @Heutel:2014]. Donors may feel less inclined to contribute to NGOs receiving funds from agencies like USAID because such funding can signal a lack of independence from the government. Organizations that receive substantial funding from their home governments frequently avoid programming that would question the donor government even if such programming is in line with their mission [@Stroup:2012]. In contrast, individual donors who see that organizations are privately funded may feel that they can also contribute and help and that their marginal donation would make a noticeable difference.

@@

> **H~3a~**: If donors learn that legally besieged NGOs receive substantial funding from government sources, then individual donors will be less likely to donate to them.
>
> **H~3b~**: If donors learn that legally besieged NGOs receive substantial funding from government sources, then individual donors will donate less to them.

Importantly, when testing these hypotheses, we do not assume that individual donors have perfect knowledge about an NGO or the country an NGO operates in. Donors do not always know about anti-NGO laws in host countries or who funds an NGO when they make their contributions. Moreover, the average donor will not know exactly what "legal crackdown" entails for any given organization (i.e. does it mean an NGO was expelled, or that its assets were frozen, or that it received a fine?). For the sake of this experiment, this ambiguous definition is not a central concern as the heuristics we explore here are explicitly simplified shortcuts to donor information-gathering. Knowing that an organization works with human rights issues (even if donors are not familiar with specific kinds of rights) should be enough to trigger the issue-based deservingess heuristic. Similarly, knowing that an NGO faces legal trouble should be enough to trigger the structural deservingness heuristic. Moreover, while many NGOs and nonprofits make information about their funding sources easily accessible through annual reports, donors often do not take the time to track down that information. In each of these hypotheses, we assume that donors will not spend additional effort to research the exact details concerning crackdown, issue area, and funding sources. Instead, donors rely on signals in basic information about the organization to shape their preferences and behavior.

We also posit that these signals work interactively to shape donor behavior and that combinations of characteristics will have different effects. For instance, the effect of crackdown will be different depending if an organization focuses on humanitarian assistance or human rights, or if it receives government funding or not. Because it is unclear a priori what these interactive effects will be, we test nested versions of the hypotheses in our analysis.

@@

To measure the effect of structural and organization-level heuristics on individual donor preferences, we use an experiment to vary the different signals donors are exposed to. Comparing these effects allows us to measure the relative strength of these heuristics, and ultimately help NGOs tailor their fundraising strategies for private individuals. Prior to launching the experiment, we preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at [https://osf.io/dx973/](https://osf.io/dx973/). We made two minor deviations from the original preregistration, described in detail in the online appendix.[^deviations] The online appendix also includes the full text of the survey experiment, as well as details about our sample demographics, balance across experimental conditions, and Bayesian priors. The analysis is fully reproducible using code and data available at ANONYMIZED.

@@

We test these hypotheses with a vignette-based survey experiment fielded through Amazon's Mechanical Turk.[^mturk] Our target population is the portion of Americans hypothetically willing to donate money for human rights and humanitarian work abroad. Our convenience sample (March 2018, N = `r nrow(results)`) generally approximates the characteristics of our target population, since it is younger, more educated, wealthier, and more likely to donate to charities than nationally representative samples. Additionally, the majority of the sample (≈90%) feels favorably toward human rights, humanitarian, and development NGOs.

@@

We presented participants with a short paragraph with three manipulated frames, each highlighting different factors that might influence patterns of private philanthropy: (1) crackdown/no crackdown, (2) humanitarian assistance/human rights, and (3) government/private funding. We used the International Rescue Committee (IRC) as our example NGO because its work easily applies to each possible frame. IRC provides humanitarian assistance and engages in human rights advocacy for refugees, more than a quarter of IRC's income comes from both government grants and private donations, and it works in countries with and without anti-NGO laws. We use a 2×2×2 between-subject factorial design with participants randomly assigned to one of eight versions of the following vignette:[^consort]

@@

We measured two outcomes: (1) how likely participants would be to donate to IRC (measured with a 5-point scale ranging from "Extremely likely" to "Extremely unlikely"), and (2) how much participants would hypothetically donate to IRC if they had an extra $100. We collapsed the likelihood scale into a binary variable measuring whether the participant is likely ("Extremely likely" and "Somewhat likely") or not likely (all other responses) to donate.[^collapse] We also included an exploratory free response question asking respondents to justify their choice.

Each of the varied treatments correspond to our hypotheses: crackdown vs. no crackdown, humanitarian assistance vs. human rights, and government vs. nongovernmental funding. Participants might misinterpret the vignette and assume that IRC's host governments fund the organization rather than its home government. In either case, knowing about government funding should still serve as a signal of the organization's deservingness. Similarly, there could be confusion with phrase "private funding", but since we use it as the opposite of government funding, it should still act as a heuristic. Free responses tend to confirm the correct interpretation of the vignette, and many respondents explicitly justified their lack of support for IRC because of government funding, as expected ("I see that a large portion of its funding comes from government donors, so I feel it doesn't need me as much."[^6216778])

@@

@tbl:hypotheses provides a summary of how our results map onto our hypotheses, both individually and nested within each other.

\stgroup

\renewcommand*{\arraystretch}{1.35}
